Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-10-24 paper/main.tex
complete draft of responses to all reviewer comments!

Designing and building procedures and tools for mapping out knowledge touches
on deep questions about what it means to learn. For example, how do we acquire
conceptual knowledge? Memorizing course lectures or textbook chapters by rote
can lead to the superficial \textit{appearance} of understanding the underlying
content, but achieving true conceptual understanding seems to require something
deeper and richer. Does conceptual understanding entail connecting newly
acquired information to the scaffolding of one's existing knowledge or
experience~\citep{BlayEtal06,CaraMaho03, ConsEtal16, DeacEtal04, SimoEtal04}?
Or weaving a lecture's atomic elements (e.g., its component words) into a
structured network that describes how those individual elements are
related~\citep{LeeChen22}? Conceptual understanding could also involve building
a mental model that transcends the meanings of those individual atomic elements
by reflecting the deeper meaning underlying the gestalt whole~\citep{Kint70,
Macl05, ScotEtal07}.

The difference between ``understanding'' and ``memorizing,'' as framed by
researchers in education, cognitive psychology, and cognitive neuroscience
(e.g.,~\citealp{Kato40, Gall00, ScotEtal07, HallGree08, Macl05}), has profound
analogs in the fields of natural language processing and natural language
understanding. For example, considering the raw contents of a document (e.g.,
its constituent symbols, letters, and words) might provide some clues as to
what the document is about, just as memorizing a passage might provide some
ability to answer simple questions about it. However, text embedding models
(e.g.,~\citealp{LandDuma97, DeerEtal90, BleiEtal03, BleiLaff06, MikoEtal13a,
CerEtal18, BrowEtal20}) also attempt to capture the deeper meaning
\textit{underlying} those atomic elements. These models consider not only the
co-occurrences of those elements within and across documents, but (in many
cases) also patterns in how those elements appear across different scales
(e.g., sentences, paragraphs, chapters, etc.), the temporal and grammatical
properties of the elements, and other high-level characteristics of how they
are used~\citep{Mann20, Mann21a}. According to these models, the deep
conceptual meaning of a document may be captured by a feature vector in a
high-dimensional representation space, wherein nearby vectors reflect
conceptually related documents. A model that succeeds at capturing an analogue
of ``understanding'' is able to assign nearby feature vectors to two
conceptually related documents, \textit{even when the specific words contained
in those documents have very little overlap}.

@@

    \caption{\textbf{Estimating moment-by-moment knowledge acquisition.}
    \textbf{A. Moment-by-moment knowledge about the \textit{Four Fundamental
    Forces}}. Each trace displays the weighted proportion of correctly answered
    questions about the content reflected in each moment of the lecture (see
    \nameref{subsec:traces}), using responses from a single quiz (color). The
    traces are averaged across participants. \textbf{B. Average estimated
    knowledge about the \textbf{Four Fundamental Forces}.} Each bar displays
    the across-timepoint average knowledge, estimated using the responses to
    one quiz's questions. \textbf{C. Moment-by-moment knowledge about the
    \textit{Birth of Stars}}. The panel is in the same format as Panel A, but
    here the knowledge estimates are for the moment-by-moment content of the
    \textit{Birth of Stars} lecture. \textbf{D. Average estimated knowledge
    about the \textit{Birth of Stars}.} The panel is in the same format as
    Panel B, but here the knowledge estimates are for the content of the
    \textit{Birth of Stars} lecture. All panels: error ribbons and error bars
    denote 95\% confidence intervals, estimated across participants.}

@@

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should
carry some predictive information about whether the participant is likely to
answer that question correctly or incorrectly. We developed a statistical
approach to test this claim. For each question, in turn, we used
Equation~\ref{eqn:prop} to estimate each participant's knowledge at the given
question's embedding space coordinate, using all \textit{other} questions that
participant answered on the same quiz. For each quiz, we grouped these
estimates into two distributions: one for the estimated knowledge at the
coordinates of \textit{correctly} answered questions, and another for the
estimated knowledge at the coordinates of \textit{incorrectly} answered
questions (Fig.~\ref{fig:predictions}). We then used independent samples
$t$-tests to compare the means of these distributions of estimated knowledge.

@@

    \caption{\textbf{Estimating knowledge at the embedding coordinates of
    held-out questions.} Separately for each quiz (panel), we plot the
    distributions of predicted knowledge at the embedding coordinates of each
    held-out correctly (blue) or incorrectly (red) answered question. The
    $t$-tests reported in each panel are between the distributions of estimated
    knowledge at the coordinates of correctly versus incorrectly answered
    held-out questions.}

@@

For the initial quizzes participants took (prior to watching either lecture),
participants' estimated knowledge tended to be low overall, and relatively
unstructured (Fig.~\ref{fig:predictions}, left panel). When we held out
individual questions and estimated their knowledge at the held-out questions'
embedding coordinates, we found no reliable differences in the estimates when
the held-out question had been correctly versus incorrectly answered ($t(633) =
0.577,~p = 0.564$). This reflects a floor effect: when knowledge is low
everywhere, there is little signal to differentiate between what is known
versus unknown. After watching the first lecture, estimated knowledge for
held-out correctly answered questions (from the second quiz;
Fig.~\ref{fig:predictions}, middle panel) exhibited a positive shift relative
to held-out incorrectly answered questions ($t(633) = 3.961,~p < 0.001$). This
second quiz provides the maximally sensitive test for our knowledge
predictions, since (if knowledge is estimated accurately) participants' Quiz~2
responses should demonstrate specific knowledge about Lecture~1 content, but
knowledge about Lecture~2 and general physics concepts should be roughly
unchanged from before they watched Lecture~1. After watching the second
lecture, estimated knowledge (from the third quiz; Fig.~\ref{fig:predictions},
right panel) for \textit{all} questions exhibited a positive shift. However,
the estimated knowledge for held-out correctly answered questions remained
greater than that for held-out incorrectly answered questions ($t(628) =
2.045,~p = 0.041$). This third contrast reflects a ceiling effect: when
knowledge is relatively high everywhere, the signal differentiating what is
known versus unknown is relatively weak. Taken together, this set of analyses
demonstrates that our knowledge prediction framework is most informative when
participants exhibit variability in their knowledge of the content captured by
the text embedding model.

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{figs/knowledge-smoothness}

    \caption{\textbf{Caption title.} Caption content.}

@@

@@

@@

@@

We then hand-created 39 multiple-choice questions: 15 about the conceptual
content of \textit{Four Fundamental Forces} (i.e., Lecture~1), 15 about the
conceptual content of \textit{Birth of Stars} (i.e., Lecture~2), and 9
questions that tested for general conceptual knowledge about basic physics
(covering material that was not presented in either video). One of our group's
undergraduate research assistants worked alongside a rotating Masters student
to develop this set of questions (these researchers are acknowledged in our
paper for their contribution, although they did not meet the criteria for
authorship discussed with all team members at the start of the project, as
determined by J.R.M.) The senior author (J.R.M.) tasked the pair of researchers
with coming up with ``15 conceptual questions about each lecture, along with 9
additional questions about general physics knowledge.'' To help broaden the set
of lecture-specific questions, the researchers were further instructed to work
through each lecture in small segments, identify what each segment was
``about'' conceptually, and then write a question about that concept. The
general physics questions were drawn from the researchers' coursework along
with internet searches and brainstorming with the project team and other
members of J.R.M.'s lab. The final set of questions (and response options) was
reviewed and approved by J.R.M.

@@

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. The transcripts
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the video (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). We defined a sliding window
length of (up to) $w = 30$ transcript lines, and assigned each window a
timestamp corresponding to the midpoint between the timestamps for its first
and last lines. This $w$ parameter was chosen to match the same number of words
per sliding window (rounded to the nearest whole word, and after preprocessing)
as the sliding windows we defined in our prior work~\citep{HeusEtal21}. The
sliding windows contained an average of 73.8 words and lasted for an average of
62.22 seconds. After splitting the transcripts into sliding windows, we carried
out several standard preprocessing steps. First, we removed all punctuation,
replaced numbers with their word equivalents (e.g., ``1'' with ``one''), and
converted all text to lowercase. Next, we removed all stop words, as defined by
the Natural Language Toolkit~\citep[NLTK; ]{BirdEtal09} English stop word list,
augmented with additional words following~\citep{BoydMimn14}: ``actual,''
``actually,'' ``also,'' ``bit,'' ``could,'' ``e,'' ``even,'' ``first,''
``follow,'' ``following,'' ``four,'' ``let,'' ``like,'' ``mc,'' ``really,'',
``saw,'' ``see,'' ``seen,'' ``thing,'' and ``two.''

These sliding windows ramped up and down in length at the beginning and end of
each transcript, respectively. In other words, each transcript's first sliding
window covered only its first line, the second sliding window covered the first
two lines, and so on. This ensured that each line from the transcripts appeared
in the same number ($w$) of sliding windows. After performing various standard
text preprocessing (e.g., normalizing case, lemmatizing, removing punctuation
and stop-words), we treated the text from each sliding window as a single
``document,'' and combined these documents across the two videos' windows to
create a single training corpus for the topic model.

@@

>>>>

Designing and building procedures and tools for mapping out knowledge touches
on deep questions about what it means to learn. For example, how do we acquire
conceptual knowledge? Memorizing course lectures or textbook chapters by rote
can lead to the superficial \textit{appearance} of understanding the underlying
content, but achieving true conceptual understanding seems to require something
deeper and richer. Does conceptual understanding entail connecting newly
acquired information to the scaffolding of one's existing knowledge or
experience~\citep{BlayEtal06,CaraMaho03, ConsEtal16, DeacEtal04, SimoEtal04,
HuebWill18}? Or weaving a lecture's atomic elements (e.g., its component words)
into a structured network that describes how those individual elements are
related~\citep{LeeChen22, vanPEtal21}? Conceptual understanding could also
involve building a mental model that transcends the meanings of those
individual atomic elements by reflecting the deeper meaning underlying the
gestalt whole~\citep{Kint70, Macl05, ScotEtal07, TulcEtal23}.

The difference between ``understanding'' and ``memorizing,'' as framed by
researchers in education, cognitive psychology, and cognitive neuroscience
(e.g.,~\citealp{Kato40, Gall00, ScotEtal07, HallGree08, Macl05}), has profound
analogs in the fields of natural language processing and natural language
understanding. For example, considering the raw contents of a document (e.g.,
its constituent symbols, letters, and words) might provide some clues as to
what the document is about, just as memorizing a passage might provide some
ability to answer simple questions about it. However, text embedding models
(e.g.,~\citealp{LandDuma97, DeerEtal90, BleiEtal03, BleiLaff06, MikoEtal13a,
CerEtal18, BrowEtal20, ViswEtal17}) also attempt to capture the deeper meaning
\textit{underlying} those atomic elements. These models consider not only the
co-occurrences of those elements within and across documents, but (in many
cases) also patterns in how those elements appear across different scales
(e.g., sentences, paragraphs, chapters, etc.), the temporal and grammatical
properties of the elements, and other high-level characteristics of how they
are used~\citep{Mann20, Mann21a}. According to these models, the deep
conceptual meaning of a document may be captured by a feature vector in a
high-dimensional representation space, wherein nearby vectors reflect
conceptually related documents. A model that succeeds at capturing an analogue
of ``understanding'' is able to assign nearby feature vectors to two
conceptually related documents, \textit{even when the specific words contained
in those documents have very little overlap}. In this way, ``concepts'' are
defined implicitly by the model's geometry~\citep[e.g., how the embedding coordinate
of a given word or document relates to the coordinates of other text
embeddings;][]{PianHill22}.

@@

    \caption{\textbf{Estimating knowledge about the content presented at each
    moment of each lecture.} \textbf{A. Knowledge about the time-varying
    content of \textit{Four Fundamental Forces}}. Each trace displays the
    weighted proportion of correctly answered questions about the content
    reflected in each moment of the lecture (see \nameref{subsec:traces}),
    using responses from a single quiz (color). The traces are averaged across
    participants. \textbf{B. Average estimated knowledge about \textbf{Four
    Fundamental Forces}.} Each bar displays the across-timepoint average
    knowledge, estimated using the responses to one quiz's questions.
    \textbf{C. Knowledge about the time-varying content of \textit{Birth of
    Stars}}. The panel is in the same format as Panel A, but here the knowledge
    estimates are for the moment-by-moment content of the \textit{Birth of
    Stars} lecture. \textbf{D. Average estimated knowledge about \textit{Birth
    of Stars}.} The panel is in the same format as Panel B, but here the
    knowledge estimates are for the content of the \textit{Birth of Stars}
    lecture. All panels: error ribbons and error bars denote 95\% confidence
    intervals, estimated across participants.}

@@

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should
carry some predictive information about whether the participant is likely to
answer that question correctly or incorrectly. We developed a statistical
approach to test this claim. For each question, in turn, we used
Equation~\ref{eqn:prop} to estimate each participant's knowledge at the given
question's embedding space coordinate, using all \textit{other} questions that
participant answered on the same quiz. For each quiz, we grouped these
estimates into two distributions: one for the estimated knowledge at the
coordinates of \textit{correctly} answered questions, and another for the
estimated knowledge at the coordinates of \textit{incorrectly} answered
questions (Fig.~\ref{fig:predictions}). We then used Mann-Whitney $U$-tests to
compare the means of these distributions of estimated knowledge.

@@

    \caption{\textbf{Estimating knowledge at the embedding coordinates of
    held-out questions.} Separately for each quiz (panel), we plot the
    distributions of predicted knowledge at the embedding coordinates of each
    held-out correctly (blue) or incorrectly (red) answered question. The
    Mann-Whitney $U$-tests reported in each panel are between the distributions
    of estimated knowledge at the coordinates of correctly versus incorrectly
    answered held-out questions. In the top row (``All questions''), we used
    all quiz questions (from each quiz, for each participant) except one to
    estimate knowledge at the held-out question's embedding coordinate. In the
    middle rows (``Across-lecture''), we used all questions about one lecture
    to estimate knowledge at the embedding coordinate of a held-out question
    about the \textit{other} lecture. In the bottom row (``Within-lecture''),
    we used all but one question about one lecture to estimate knowledge at the
    embedding coordinate of a held-out question about the \textit{same}
    lecture. We repeated each of these analyses using all possible held-out
    questions for each quiz and participant.}

@@

For the initial quizzes participants took (prior to watching either lecture),
participants' estimated knowledge tended to be low overall, and relatively
unstructured (Fig.~\ref{fig:predictions}, left column). When we held out
individual questions and estimated their knowledge at the held-out questions'
embedding coordinates, we found no reliable differences in the estimates when
the held-out question had been correctly versus incorrectly answered. This
``null'' effect persisted when we used \textit{all} of the Quiz 1 questions
from a given participant to predict a held-out question (``All questions''; $U
= 50587, p = 0.723$), when we used questions from one lecture to predict
knowledge at the embedding coordinate of a held-out question about the
\textit{other} lecture (``Across-lecture''; predicting knowledge for held-out
\textit{Four Fundamental Forces Questions} using \textit{Birth of Stars}
questions: $U = 8244, p = 0.184$; predicting knowledge for held-out
\textit{Birth of Stars} questions: $U = 8202.5, p = 0.161$), and when we used
questions from one lecture to predict knowledge at the embedding coordinate of
a held-out question about the \textit{same} lecture (``Within-lecture'';
\textit{Four Fundamental Forces}: $U = 7681.5, p = 0.746$; \textit{Birth of
Stars}: $U = 8125, p = 0.204$). We believe that this reflects a floor effect:
when knowledge is low everywhere, there is little signal to differentiate
between what is known versus unknown.

After watching \textit{Four Fundamental Forces}, estimated knowledge for
held-out correctly answered questions (from the second quiz;
Fig.~\ref{fig:predictions}, middle column) exhibited a positive shift relative
to held-out incorrectly answered questions. This held when we included all
questions in the analysis ($U = 58332, p < 0.001$), when we predicted knowledge
across-lectures (\textit{Four Fundamental Forces}: $U = 6749.5, p = 0.014$;
\textit{Birth of Stars}: $U = 8480, p = 0.016$), and when we predicted
knowledge at the embedding coordinates of held-out \textit{Four Fundamental
Forces} questions using other \textit{Four Fundamental Forces} questions from
the same quiz and participant ($U = 7224, p < 0.001$). This difference did
\textit{not} hold for within-lecture predictions of \textit{Birth of Stars}
knowledge ($U = 7419, p = 0.739$). Again, we suggest that this might reflect a
floor effect whereby knowledge about the content of the \textit{Birth of Stars}
material is relatively low everywhere in that region of text embedding space.

Finally, after watching \textit{Birth of Stars}, estimated knowledge for
held-out correctly answered questions (from the third quiz;
Fig.~\ref{fig:predictions}, right column) was higher for held-out correctly
answered questions than for held-out incorrectly answered questions. This held
when we included all questions in the analysis ($U = 38279, p = 0.022$), when
we carried out across-lecture predictions (\textit{Four Fundamental Forces}: $U
= 6684.5, p = 0.032$; \textit{Birth of Stars}: $U = 6414.5, p = 0.002$), and
and when we carried out within-lecture predictions of held-out \textit{Birth of
Stars} questions using other \textit{Birth of Stars} questions from the same
quiz and participant ($U = 6126, p = 0.006$). However, we found the
\textit{opposite} effect when we carried out within-lecture predictions of
held-out \textit{Four Fundamental Forces} questions using other \textit{Four
Fundamental Forces} questions from the same quiz and participant ($U = 6734, p
= 0.027$). Specifically, held-out correctly answered Quiz 3 questions about
\textit{Four Fundamental Forces} had reliably \textit{lower} estimated
knowledge than held-out incorrectly answered questions. Speculatively, we
suggest that this may reflect participants forgetting some of the \textit{Four
Fundamental Forces} content. If this forgetting happens in a relatively
``random'' way (with respect to spatial distance within the text embedding
space), then it could explain why some held-out questions about \textit{Four
Fundamental Forces} were answered incorrectly, even if questions at nearby
coordinates (i.e., about similar content) were answered correctly. This might
lead our approach to over-estimate knowledge for held-out questions about
``forgotten'' knowledge that particpants answered incorrectly.

That the knowledge estimates derived from the text embedding space reliably
distinguish between held-out correctly versus incorrectly answered questions
(Fig.~\ref{fig:predictions}) suggests that the text embedding space bears at
least some relationship to participants' knowledge. But what does that
relationship look like as we move through the embedding space? For example,
suppose we know that a participant answers a question (at embedding coordinate
$X$) correctly. As we move away from $X$ in the embedding space, how does quiz
performance ``fall off'' with distance? Or, suppose the participant instead
answered that same question \textit{in}correctly. Again, as we move away from
$X$ in the embedding space, how does our confidence that the participant does
\textit{not} know about the content change with distance? We reasoned that,
assuming our space is capturing something about how participants actually
organize their knowledge, conceptual knowledge right around $X$ should be similar
to the participant's knowledge of the content at $X$. And at another extreme, at
some distance (after moving sufficiently far away from $X$), our guesses about
what participants know (based on their response to the question at location $X$)
should be no better than guessing based on their overall proportion of
correctly answered questions—i.e., if $Y$ is very far away from $X$, all we can do
with the participant's response to $X$ is guess that ``their performance on quiz
questions about $Y$ is about equal to their average performance on quiz questions
about any material.''

\begin{figure}[t]
    \centering
    \includegraphics[width=0.8\textwidth]{figs/knowledge-smoothness}

    \caption{\textbf{Quiz performance falls off gradually in text embedding
    space.} \textbf{A. Performance versus distance.} For each participant, for
    each correctly answered question (blue) or incorrectly answered question
    (red), we computed the proportion of correctly answered questions within a
    given distance of that question's embedding coordinate. We repeated this
    analysis for all questions and participants, and separately for each quiz
    (column). The black lines denote the average proportion correct across
    \textit{all} questions included in the analysis at the given distance.
    \textbf{B. Maximum distance for which performance is reliably different
    from the average.} We used a bootstrap procedure
    (see~\nameref{subsec:smoothness}) to estimate the point at which the blue
    and red lines in Panel A reliably diverged from the black line. We repeated
    this analysis separately for correctly and incorrectly answered questions
    from each quiz. \textbf{All panels.} Error ribbons denote
    bootstrap-estimated 95\% confidence intervals.}

@@

With these ideas in mind, we asked: conditioned on answering a question
correctly, what proportion of all questions (within some radius, $r$, of that
question's embedding coordinate) were answered correctly? We plotted this
proportion as a function of $r$. Similarly, we could ask, conditioned on
answering a question incorrectly, how the proportion of correct responses
changed with $r$. As shown in Figure~\ref{fig:smoothness}, we found that quiz
performance falls off smoothly with distance, and the ``rate'' of the falloff
does not appear to change across the different quizzes, as measured by the
distance at which performance becomes statistically indistinguishable from a
simple proportion correct score (see~\nameref{subsec:smoothness}). This
suggests that, at least within the region of text embedding space covered by
the questions our participants answered (and as characterized using our topic
model), the rate at which knowledge changes with distance is relatively
constant, even as participants' overall level of knowledge varies across
quizzes or regions of the embedding space.

@@

Another application for large language models that does \textit{not} require
explicitly modeling the content of individual lectures or questions is to
leverage the models' ability to generate text. For example, generative text
models like ChatGPT~\citep{ChatGPT} and LLaMa~\citep{TouvEtal23} are already
being used to build a new generation of interactive tutoring
systems~\citep[e.g.,][]{MannEtal23b}.  Unlike the approach we have taken here,
these systems do not explicitly model what learners know, or how their knowledge 
changes over time with training.  One could imagine building a hybrid system
that combines the best of both worlds: a large language model that can
\textit{generate} text, combined with a smaller model that can \textit{infer}
what learners know and how their knowledge changes over time.  Such a hybrid
system could potentially be used to build a new generation of interactive
tutoring systems that are able to adapt to learners' needs in real time, and
that are able to provide more nuanced feedback about what learners know and
what they do not know.

@@

We view our approach as occupying a sort of ``sweet spot,'' between much larger
language models and simple word matching-based approaches, that enables us to
capture the relevant conceptual content of course materials at an appropriate
semantic scale. Our approach enables us to accurately and consistently identify
each question's content in a way that also matches up with what is presented in
the lectures. In turn, this enables us to construct accurate predictions about
participants' knowledge of the conceptual content tested by held-out questions
(Fig.~\ref{fig:predictions}).

One limitation of our approach is that topic models contain no explicit
internal representations of more complex aspects of ``knowledge,'' like
knowledge graphs, dependencies or associations between concepts, causality, and
so on. These representations might (in principle) be added as extensions to our
approach to more accurately and precisely capture, characterize, and track
learners' knowledge. However, modeling these aspects of knowledge will likely
require substantial additional research effort.

@@

We then hand-created 39 multiple-choice questions: 15 about the conceptual
content of \textit{Four Fundamental Forces} (i.e., Lecture~1), 15 about the
conceptual content of \textit{Birth of Stars} (i.e., Lecture~2), and 9
questions that tested for general conceptual knowledge about basic physics
(covering material that was not presented in either video). One of our group's
undergraduate research assistants worked alongside a rotating Masters student
to develop this set of questions (these researchers are acknowledged in our
paper for their contribution, although they did not meet the criteria for
authorship discussed with all team members at the start of the project, as
determined by J.R.M.) The senior author (J.R.M.) tasked the pair of researchers
with coming up with ``15 conceptual questions about each lecture, along with 9
additional questions about general physics knowledge.'' To help broaden the set
of lecture-specific questions, the researchers were further instructed to work
through each lecture in small segments, identify what each segment was
``about'' conceptually, and then write a question about that concept. The
general physics questions were drawn from the researchers' coursework along
with internet searches and brainstorming with the project team and other
members of J.R.M.'s lab. The final set of questions (and response options) was
reviewed and approved by J.R.M. before we collected or analyzed the text or
experimental data.

@@

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. The transcripts
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the video (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). We defined a sliding window
length of (up to) $w = 30$ transcript lines, and assigned each window a
timestamp corresponding to the midpoint between the timestamps for its first
and last lines. This $w$ parameter was chosen to match the same number of words
per sliding window (rounded to the nearest whole word, and before
preprocessing) as the sliding windows we defined in our prior
work~\citep{HeusEtal21} (i.e., 185 words per sliding window).

These sliding windows ramped up and down in length at the beginning and end of
each transcript, respectively. In other words, each transcript's first sliding
window covered only its first line, the second sliding window covered the first
two lines, and so on. This ensured that each line from the transcripts appeared
in the same number ($w$) of sliding windows. We next performed a series of
standard text preprocessing steps: normalizing case, lemmatizing, removing
punctuation and removing stop-words. We constructed our corpus of stop words by
augmenting the Natural Language Toolkit~\citep[NLTK; ]{BirdEtal09} English stop
word list with the following additional words, selected using the stop word
approach suggested by~\citep{BoydMimn14}: ``actual,'' ``actually,'' ``also,''
``bit,'' ``could,'' ``e,'' ``even,'' ``first,'' ``follow,'' ``following,''
``four,'' ``let,'' ``like,'' ``mc,'' ``really,'', ``saw,'' ``see,'' ``seen,''
``thing,'' and ``two.'' This yielded sliding windows with an average of 73.8
remaining words, and lasting for an average of 62.22~seconds. We treated the
text from each sliding window as a single ``document,'' and combined these
documents across the two videos' windows to create a single training corpus for
the topic model.

@@

\subsubsection*{Estimating the ``smoothness'' of knowledge}\label{subsec:smoothness}

In the analysis reported in Figure~\ref{fig:smoothness}A, we show how
participants' quiz performance changes as a function of distance to a given
correctly or incorrectly answered reference question. We used a bootstrap-based
approach to estimate the maximum distances over which these proportions
of correctly answered questions could be reliably distinguished from participants'
overall average proportion of correctly answered questions.

In our bootstrap procedure, we ran 10,000 iterations to estimate the
relationship between participants' performance and the distance to a given
reference question. For each of these iterations, for every individual quiz
($q$), we first determined the across-participants average ``simple''
proportion correct and its 95\% confidence interval. This interval was
established by repeatedly (1,000 times) subsampling participants with
replacement, computing the mean ``simple'' proportion correct for each
subsample, and then deriving the 2.5\textsuperscript{th} and
97.5\textsuperscript{th} percentiles from the distribution of these subsample
means. We used this interval as our benchmark for determining whether the
proportion of correctly answered questions for a given subset of questions was
reliably different (at the $p < 0.05$ significance level) from the average
proportion correct across all questions.

Next, for each participant, we examined all 15 questions they answered on quiz
$q$. We treated each question as the ``reference question'' in turn. Around
this reference, we constructed a series of 15-dimensional spheres (starting
with a radius of 0), where each successive circle had a radius of 0.01
(correlation distance) greater than its predecessor. Within each of these
spheres, we calculated the proportion of questions answered correctly by the
participant. The per-radius proportion correct values were then averaged across
both categories of ``reference questions'': those answered correctly and those
answered incorrectly. This yielded two distinct proportion-correct values for
each binned distance (radius) for a specific participant and quiz. From these,
we established the average proportion correct within each radius for both
categories of reference questions. Finally, we identified the minimum binned
distance from the correctly answered reference questions for which the average
proportion correct intersected the 95\% confidence interval of the simple
average proportion correct computed earlier. (A parallel analysis was conducted
for the incorrectly answered reference questions.) We display the resulting
distance estimates, for each quiz and reference question status, in
Figure~\ref{fig:smoothness}B.
